---
title: "Stats 506, F20, Group Project"
author: "Group 6: Erin Susan Cikanek, Suppapat Korsurat, Kyle William Schulz"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: paper
    df_print: paged
    toc: TRUE
---

```{r, include = FALSE, echo = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(knitr)
```

### Introduction

### Data 

### Method

### Core Analysis {.tabset}
#### Python
```{r, eval = FALSE}
"Kyle's Code"
```

```{r, eval = FALSE}
"When you want to show only code, but prevent this chunck to run."
```

```{r, echo = FALSE}
print("When you want this chunck to run, but don't want to show the code.")
```

#### Stata
```{r, eval = FALSE}
"Erin's Code"
```

```{r, eval = FALSE}
"When you want to show only code, but prevent this chunck to run."
```

```{r, echo = FALSE}
print("When you want this chunck to run, but don't want to show the code.")
```

#### R
The library `splines` is required for implementing splines by using R.
```{r}
library(splines)
```

```{r, include = FALSE}
data <- read.csv("/Users/kevin/506FA20/Stats506_Project/Dataset/data.csv")
```

```{r scatter_wage_age, include = FALSE}
wage_age <- function(line = FALSE, poly = 1, formula_wage_age = y ~ x){
  plot_title <- "Scatter Plot between Wage and Age"
  
  if(poly != 1){
    plot_title <- paste0("Polynomial degree ", poly)
  }
  
  plot_wage_age <- ggplot(data, aes(x = age, y = wage)) + 
    geom_point(color = "darkblue") +
    theme_bw() +
    labs(title = plot_title, x = "Age", y = "Wage")
  
  if(line == TRUE){
    plot_wage_age + 
      geom_smooth(method = "lm", formula = formula_wage_age, color = "yellow")
  } else {
    plot_wage_age
  }
}
```

First, considering the linear regression.
```{r}
model <- lm(wage ~ age + education + year, data = data)
summary(model)
```

The $R^2$ is 0.2619, which is pretty low. Consider the scatter plot between 
`Wage` and `Age`.

```{r, echo = FALSE, fig.cap = fig_s1}
wage_age()
fig_s1 <- paste0("Figure 3.1 Scatter plot between Wage and Age")
```

The scatter plot show that the relationship between these two variables are not linear. Hence, we will try various types of spline.

##### Step Function
Consider applying the step function on `Age`. 
```{r}
model_cut <- lm(wage ~ cut(age, 4) + education + year, data = data)
summary(model_cut)
```

The $R^2$ is 0.2828, which improved from the previous model. The plot below 
is a scatterplot between `Wage` and `Age`, also the yellow line represents 
the step function.

```{r, echo = FALSE, fig.cap = fig_s2}
wage_age(TRUE, 1, y ~ cut(x,4))
fig_s2 <- paste0("Figure 3.2 Scatter plot between Wage and Age with the
                 step function.")
```

##### Polynomial Regression
Consider the various number for the degree in the polynomial regression. The plots below are the result from the fitting polynomial regression.
```{r, echo = FALSE, warning = FALSE}
degree_poly <- c(2,3,5,20)

for(i in 1:length(degree_poly)){
  poly_d <- degree_poly[i]
  print(wage_age(TRUE, poly_d, y ~ poly(x, poly_d)))
}

```

```{r, echo = FALSE}
poly_r2 <- c()

for(i in 2:20){
  r2_p <-summary(lm(wage ~ poly(age, i) + education + year, 
                   data = data))$r.squared
  poly_r2 <- c(poly_r2, r2_p)
}

poly_r2 <- data.frame(degree = 2:20, R2 = poly_r2)
colnames(poly_r2) <- c("Degree of the age polynomial", "R-Squared")
poly_r2
```

Even the higher degree give the higher $R^2$, the `overfitting` problem may be occured. Hence, polynomial regression with degree 3 would be appropriate.

```{r}
model_poly <- lm(wage ~ poly(age, 3) + education + year, data = data)
summary(model_poly)
```

The $R^2$ is 0.2909, which improved from all previous models.

##### Basis Spline and Natural Spline

### Summary
### Discussion
### Reference


